Data Management and Domain Adjusting Strategies for Implementing Parallel SPH Method in JPUE Simulation

### 2016-01-08 Paper for submission.tex
figure 9

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra}
\IEEEauthorblockA{Department of MAE \\%Mechanical and Aerospace Engineering
 The University at Buffalo\\
 Buffalo, New York 14260\\
Email: zhixuanc@buffalo.edu, \\ abani@buffalo.edu}
\and
\IEEEauthorblockN{Matthew Jones, Jaroslaw Zola}
\IEEEauthorblockA{Center for Computational Research, \\
Department of CSE \\%Computer Science and Engineering
The University at Buffalo\\
Buffalo, New York 14260\\
Email: jonesm@buffalo.edu, \\ jzola@buffalo.edu}
%\and
%\IEEEauthorblockN{Jaroslaw Zola}
%\IEEEauthorblockA{The University at Buffalo\\
%Department of Computer Science and Engineering\\
% The University at Buffalo\\
% Buffalo, New York 14260\\
%Email:jzola@buffalo.edu}
}
% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE). Background grid is used to reduce neighbors searching cost and to decompose (redecompose) the domain. Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid. As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.  
%to distinguish particles added at the same position but different simulation time step. 
Both particle and background grid are managed by hashtables which ensure quick and flexible accessing, adding and deleting of element. %Additional link list attached to the hashtable is used to handle hash conflicts.
A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance. In addition, computational domain is adjusted during simulation to reduce computational cost. 
Numerical tests show that our code has acceptable strong scalability and good weak scalability.
These strategies can be further applied to many other implementations of meshfree methods, especially those implementations that require irregular
particle distributions, adding and deleting of particles and leverage adaptive background grids.
\end{abstract}

@@

Where $\textbf{T} \subset [0,\infty)$ is the time step dimension, $\textbf{N}=\lbrace 0, 1, 2, 3...\rbrace$.

@@

\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.23]{Involved_initial}}

@@

\subfloat[Computational domain at $t=t0 + \Delta t$]{\includegraphics[scale = 0.23]{Involved_domain_adjusting}}

@@

Rm = n_gR_g + n_aR_a  \label{eq:gov-Rm}

@@

At the eruption vent, temperature $T$, velocity $\textbf{v}$, pressure $p$, initial mass fraction of solid portition $n_{g0}$ and mass discharge rate $\dot M$ is given. 

@@

Non-slip, adiabatic wall boundary condition is imposed on the wall. The pressure of the surrounding atmosphere is specified on the pressure boundaries. 

@@

Free surface flows are inherently turbulent. We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume. This will result in a filtered velocity for position updates and additional turbulent terms in momentum and energy equation.

@@

Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. Two types of processors -- 8 core L5630 running at 2.13GHz clock rate with 3GB memory per core on a Mellanox network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic network. Each node comprised of two sockets with several of these processors. Memory and level 3 cache are shared on each node.

@@

A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. 

@@

\subfloat[Contour of mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.4cm]{Plume_simulation}}

@@

Hashtables with external link list are adopted for accessing particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimized interval. The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation. 

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra}
\IEEEauthorblockA{Department of MAE \\%Mechanical and Aerospace Engineering
 The University at Buffalo\\
 Buffalo, New York 14260\\
Email: zhixuanc@buffalo.edu, \\ abani@buffalo.edu}
\and
\IEEEauthorblockN{Matthew Jones}
\IEEEauthorblockA{Center for Computational Research\\%Computer Science and Engineering
The University at Buffalo\\
Buffalo, New York 14260\\
Email: jonesm@buffalo.edu}
\and
\IEEEauthorblockN{Jaroslaw Zola}
\IEEEauthorblockA{Department of CSE\\
The University at Buffalo\\
Buffalo, New York 14260\\
Email:jzola@buffalo.edu}
}
% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE). Background grid is used to reduce neighbors searching cost and to decompose (redecompose) the domain. Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid. As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.  
%to distinguish particles added at the same position but different simulation time step. 
Both particle and background grid are managed by hash tables which ensure quick and flexible accessing, adding and deleting of element. %Additional link list attached to the hashtable is used to handle hash conflicts.
A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance. In addition, computational domain is adjusted during simulation to reduce computational cost. 
Numerical tests show that our code has acceptable strong scalability and good weak scalability.
These strategies can be further applied to many other implementations of meshfree methods, especially those implementations that require irregular
particle distributions, adding and deleting of particles and leverage adaptive background grids.
\end{abstract}

@@

Where $\textbf{T} \subset [0,\infty)$ is the time dimension, $\textbf{N}=\lbrace 0, 1, 2, 3...\rbrace$.

@@

\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.22]{Involved_initial}}

@@

\subfloat[Computational domain at $t=t0 + \Delta t$]{\includegraphics[scale = 0.22]{Involved_domain_adjusting}}

@@

R_m = n_gR_g + n_aR_a  \label{eq:gov-Rm}

@@

At the eruption vent, temperature $T$, velocity $\textbf{v}$, pressure $p$, initial mass fraction of gas portition $n_{g0}$ and mass discharge rate $\dot M$ is given. 

@@

Non-slip, adiabatic wall boundary condition is imposed on the wall. The pressure of the surrounding atmosphere is specified, forming pressure boundary conditions. 

@@

Free surface flows are inherently turbulent. We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume. This will result in a filtered velocity for position updating and additional turbulent terms in momentum and energy equation.

@@

Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. Two types of processors -- 8 core L5520/L5630 running at 2.27/2.13GHz clock rate with 3GB memory per core on a Mellanox Infiniband network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband network. Each node comprised of two sockets with several of these processors. Memory and level 3 cache are shared on each node.

@@

A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. Input parameters for this simulation are exactly the same as run P1 in reference \cite{suzuki2005numerical}.

@@

\subfloat[Simulation results: mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.4cm]{Plume_simulation}}

@@

Hash tables with external link list are adopted for accessing particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimized interval. The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation. 
